DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-07 paper/evaluation.tex
comparison revision done
Edited evaluation

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (Tensorflow is not compatible with the GPU).
For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configure the clock speed of each core at the maximum 1.2GHz.

@@

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting about 85\% of
the execution time.
Second, more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentail time is 50.70ms.
This means that the DeepPicar can operates
at about 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, without needing any remote computing
resources. We consider these results respectable given the complexity
of the deep neural network  and the fact that the inference operation
performed by Tensorflow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow.)

In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, is reportedly run at 30 Hz~\cite{Bojarski2016},
just a bit faster than the DeepPicar. Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TK2, later in
Section~\ref{sec:comparison}), the fact that the low-cost
Raspberry Pi 3 can achieve similar real-time control performance is
surprising.

@@

In this experiment, we investigate the scalability of performing
inference operation of DeepPicar's nueral network with respect to the
number of cores. As noted earlier, the Raspberry Pi 3 platform has
four Cortex-A53 cores and the Tensorflow 
provides a programmable mechanism to to adjust how many cores to be
used by the library. Leveraging this feature, we repeat the
same experiment described in the previous subsection with varying
number of CPU cores---from one to four.

@@

Figure~\ref{fig:perf-vs-corecnt} shows the average execution times of
the control loop as we vary the number of cores used by the
Tensorflow. As expected as we assign more cores, the average execution
time decreases---from 61.96 ms on a single core to 42.67 ms on four
cores (a 30\% improvement). However, the improvement is far from ideal
linear scaling. In particular, from 2 cores to 3 cores, the
improvement is mere 2.38ms (or 4\%). In short, we find that the
scalability of DeepPicar's deep neural network is not ideal on the
platform. We do not know whether it is due to the limitations of
Tensorflow's multicore implementation or it's the model's inherent
characteristics.

The poor scalability opens up the possibility of consolidating
multiple different tasks or different nueral network models rather
than allocating all cores for a single nueral network model. For
example, it is conceivable to use four cameras and four different
neural network models, each of which is trained seperately and
executed on a single dedicated core. Assuming we use the same network
architecture for all models, then one might exepct to achieve up to
15 Hz using one core (given 1 core can deliver 62ms average
execution time). In the next experiment, we investigate the
feasibility of such a scenario.

@@

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact to their inference
timings. In particular, we are interested in how shared resource
contention affect inference timing.

@@

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure,
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores\/DNN. For example, '4x1' means running four DNN models each of
which is assigned to run on one CPU core, whereas '2x2' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
Two bars in the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1x1), when four DNN models are co-scheduled (4x1), each model
suffers an average inference time increase of approximately 15 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled, the average inference timing is increased by
about 7 ms, or 10\%, comapred to the baseline of running one model
using two CPU cores.

@@

These increases in inference times in the co-scheduled scenarios are
expected and likely caused by contention in the shared hardware
resources such as the shared L2 cache and/or the DRAM controller.
To further analyze the source of contention, we use hardwawre
performance counters of the processor. Specifically, we measure L2
miss rates of the DNN models first in isolation and then after
co-scheduling another models. If the shared L2 cache is the primary
source of inteference, then the measured L2 miss rates would
increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
As can be see in the figure, L2 miss rates are largely unchanged
regardless whether multiple models are co-scheduled or not. This
suggests that the shared L2 cache is not the bottleneck that caused
execution time increases. In other words, DNN models appear to be not
sensitive to the shared L2 cache space. % L2 partitioning is probably
                                % not going to be useful.
Instead, we hypothesize that it is likely caused by the memory
controller---the only other major shared hardware source---where
memory requests from different CPU cores contend, which would result
in increased memory access latency. While some Intel processors
provide uncore hardware counters that can measure average memory
access latency~\cite{ye2016maracas}, we were not able to identify
whether such hardware counters exist in the BCM2837 processor of
Raspberry Pi 3 due to the lack of documentation. Instead, in the next
experiment, we use memory intensive synthetic benchmarks to test the
hypothesis.

@@

\subsection{Effect of Co-scheduling Memory Performance Hogs}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
  \caption{Effect of memory performance hogs on the DNN
    inferencing. The DNN model uses Core 0 and memory-hog co-runnes
    use the rest of the cores.}
  \label{fig:}
\end{figure}

@@

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 9.6X---even though the DNN model is running on a
dedicated core (Core0). On the other hand, the DNN model's L2
cache-miss rates are not increase as much. This suggests that
the DNN model's exeuction increase cannot be fully explained by
increased L2 cache-misses. Instead, as we hypothesized in the previous
experiment, the increased memory pressure from the co-scheduled memory
intensive benchmarks is the primary cause of the DNN model's execution
time increase. Therefore, we conclude that DeepPicar's DNN model is
more senstive to DRAM access latency than L2 cache space.

@@

First, we find that DeepPicar's computing platform,
the Raspberry Pi 3 Model B, offers adquate computing capacity to
perform real-time control of the RC car at 20 Hz frequency (or
50ms per control loop). Given the complexity of the DNN we use, we
were pleasantly suprised by this finding. The time breakdown shows that
the DNN inferencing operation, perforemd by the Tensorflow library,
dominates the execution time, which is expected.

Second, we find that scalability of Tensorflow's DNN inference 
implementation is limited: We find that using all four cores are 
only about 30\% better than using just a single core.

@@

Lastly, we however find that consolidating memory (DRAM) performance
intensive applications could jeopadize DNN performance because DNN
performance appear very sensitive to memory performance; we observe up
to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 
bandwidth intensive applications on idle cores.

>>>>

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (Tensorflow is not compatible with the Raspberry Pi's GPU).
For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configure the clock speed of each core at the maximum 1.2GHz.

@@

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting about 85\% of
the execution time.
Second, more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentail time is 50.70ms.
This means that the DeepPicar can operate
at about 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, without needing any remote computing
resources. We consider these results respectable given the complexity
of the deep neural network  and the fact that the inference operation
performed by Tensorflow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow.)

In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, is reportedly run at 30 Hz~\cite{Bojarski2016},
just a bit faster than the DeepPicar. Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that the low-cost
Raspberry Pi 3 can achieve similar real-time control performance is
surprising.

@@

In this experiment, we investigate the scalability of performing
inference operations of DeepPicar's neural network with respect to the
number of cores. As noted earlier, the Raspberry Pi 3 platform has
four Cortex-A53 cores and Tensorflow 
provides a programmable mechanism to to adjust how many cores are to be
used by the library. Leveraging this feature, we repeat the
same experiment described in the previous subsection with varying
numbers of CPU cores---from one to four.

@@

Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
the control loop as we vary the number of cores used by
Tensorflow. As expected as we assign more cores, the average execution
time decreases---from 61.96 ms on a single core to 42.67 ms on four
cores (a 30\% improvement). However, the improvement is far from ideal
linear scaling. In particular, from 2 cores to 3 cores, the
improvement is mere 2.38ms (or 4\%). In short, we find that the
scalability of DeepPicar's deep neural network is not ideal on the
platform. We do not know whether it is due to the limitations of
Tensorflow's multicore implementation or it's the model's inherent
characteristics.

The poor scalability opens up the possibility of consolidating
multiple different tasks or different neural network models rather
than allocating all cores for a single neural network model. For
example, it is conceivable to use four cameras and four different
neural network models, each of which is trained separately and
executed on a single dedicated core. Assuming we use the same network
architecture for all models, then one might expect to achieve up to
15 Hz using one core (given 1 core can deliver 62ms average
execution time). In the next experiment, we investigate the
feasibility of such a scenario.

@@

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact to their inference
timings. In particular, we are interested in how shared resource
contention affects inference timing.

@@

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores\/DNN. For example, '4x1' means running four DNN models each of
which is assigned to run on one CPU core, whereas '2x2' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
Two bars in the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1x1), when four DNN models are co-scheduled (4x1), each model
suffers an average inference time increase of approximately 15 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled, the average inference timing is increased by
about 7 ms, or 10\%, compared to the baseline of running one model
using two CPU cores.

@@

These increases in inference times in the co-scheduled scenarios are
expected and likely caused by contention in the shared hardware
resources such as the shared L2 cache and/or the DRAM controller.
To further analyze the source of contention, we use hardwawre
performance counters of the processor. Specifically, we measure L2
miss rates of the DNN models first in isolation and then after
co-scheduling another models. If the shared L2 cache is the primary
source of inteference, then the measured L2 miss rates would
increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
As can be see in the figure, L2 miss rates are largely unchanged
regardless whether multiple models are co-scheduled or not. This
suggests that the shared L2 cache is not the bottleneck that caused
execution time increases. In other words, DNN models appear not to be
sensitive to the shared L2 cache space. % L2 partitioning is probably
                                % not going to be useful.
Instead, we hypothesize that it is likely caused by the memory
controller---the only other major shared hardware source---where
memory requests from different CPU cores contend, which would result
in increased memory access latency. While some Intel processors
provide oncore hardware counters that can measure average memory
access latency~\cite{ye2016maracas}, we were not able to identify
whether such hardware counters exist in the BCM2837 processor of
Raspberry Pi 3 due to the lack of documentation. Instead, in the next
experiment, we use memory intensive synthetic benchmarks to test the
hypothesis.

@@

\subsection{Effect of Co-scheduling Memory Performance Hogs}\label{sec:eval-memhog}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
  \caption{Effect of memory performance hogs on the DNN
    inferencing. The DNN model uses Core 0 and memory-hog co-runners
    use the rest of the cores.}
  \label{fig:}
\end{figure}

@@

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 9.6X---even though the DNN model is running on a
dedicated core (Core0). On the other hand, the DNN model's L2
cache-miss rates did not increase as much. This suggests that
the DNN model's exeuction increase cannot be fully explained by
increased L2 cache-misses. Instead, as we hypothesized in the previous
experiment, the increased memory pressure from the co-scheduled memory
intensive benchmarks is the primary cause of the DNN model's execution
time increase. Therefore, we conclude that DeepPicar's DNN model is
more senstive to DRAM access latency than L2 cache space.

@@

First, we find that DeepPicar's computing platform,
the Raspberry Pi 3 Model B, offers adquate computing capacity to
perform real-time control of the RC car at 20 Hz frequency (or
50ms per control loop). Given the complexity of the DNN we use, we
were pleasantly suprised by this finding. The time breakdown shows that
the DNN inferencing operation, performed by the Tensorflow library,
dominates the execution time, which is expected.

Second, we find that scalability of Tensorflow's DNN inference 
implementation is limited: We find that using all four cores is 
only about 30\% better than using just a single core.

@@

Lastly, we find that consolidating memory (DRAM) performance
intensive applications could jeopadize DNN performance because DNN
performance appears very sensitive to memory performance; we observe up
to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 
bandwidth intensive applications on idle cores.
